The Effect of Different Writing Tasks on Linguistic Style:\\ A Case Study of the ROC Story Cloze Task

### 2017-02-07 docs/language_constraint.tex
Final version

\begin{document}
\maketitle
\begin{abstract}
A writer's style depends not just on personal traits but also on her intent and mental state.
In this paper, we show how variants of the same writing task can lead to measurable differences in writing style.
We present a case study based on 
the  {\it story cloze task} \cite{Mostafazadeh:2016},
where annotators were assigned similar writing tasks with different constraints: (1) writing an entire story, (2) adding a story ending for a given story context, and (3) adding an incoherent ending to a story.\resolved{\clinic{Many felt these 2 constraints looks like 3 constraints. clarify}}
%We show that a linear classifier, which applies only simple style features, such as sentence length and character n-grams, 
We show that a simple linear classifier informed with stylistic features is able to successfully distinguish between the three cases, without even looking at the story context.
In addition, our style-based classifier establishes a new state-of-the-art result on the story cloze challenge, substantially higher than previous results based on deep learning models. 
%Furthermore, combining our model with an LSTM-based language model, which does look at the context, leads to additional gains.
Our results demonstrate that different task framings can dramatically affect the way people write.\resolved{\clinic{Similarly to previous comment: which results does this comment address?}}

@@

\section{Introduction}
Writing style is expressed through a range of linguistic elements such as words, sentence structure, and rhetorical devices.
It is influenced by personal factors such as age \cite{Schler:2006} and gender \cite{Argamon:2003}, 
by personality traits such as agreeableness and openness  \cite{Ireland:2014b},
as well as by mental states\resolved{\clinic{Avoid using ``cognitive"?}} such as sentiment \cite{Davidov:2010}, sarcasm \cite{Tsur:2010}, and deception \cite{Feng:2012}.  
In this paper, we study the extent to which writing style is affected by the nature of the writing task the writer was asked to perform, since
different tasks likely engage different cognitive processes \cite{Campbell:2003,Banerjee:2014}.

@@

While the story cloze task was originally designed to be a story understanding challenge, 
its annotation process introduced three variants of the same writing task: writing an {\it original}, {\it right}, or {\it wrong} ending to a short story.
In this paper, we show that a linear classifier informed by stylistic features can distinguish between the different endings to a large degree, even without looking at the story context (64.5--75.6\% binary classification results).

@@

In order to further validate our method, we also directly tackle the
story cloze task \nascomment{cite the challenge?  a reader may
  not realize that there is a shared task on this}\roy{Maybe ``challenge'' was not the right term here. We basically mean the story cloze task. The shared task was the original task w/o any modification, so I am not referring to it specifically here}. 
 Adapting our classifier to the task, we obtain 72.4\% accuracy, a
 12.5\% increase over the previously reported state of the art  \cite{Salle:2016}.
We also show that the style differences captured by our model can be combined with neural language models to make a better use of the story context. 
Our final model that combines context with stylistic features achieves
75.2\%---an additional 2.8\% gain, 15.3\% better than the best
published result.

@@

The resulting stories, both {\it right} and {\it wrong}, were then individually rated for coherence and meaningfulness by additional AMT workers. \resolved{\nascomment{by
-  additional AMT workers? or someone else?}} 
Only stories rated as simultaneously coherent with a {\it right} ending and neutral with a {\it wrong} ending were selected for the task. 
It is worth noting that workers rated the stories as a whole, not only the endings.

@@

\isection{Surface Analysis of the Story Cloze Task}{Surface}
We computed several characteristics of the three types of endings: {\it original} endings (from the ROC story corpus training set), {\it right} endings and {\it wrong} endings (both from the story cloze task development set).
Our analysis  reveals several style differences between different groups. 
First, {\it original} endings are on average longer (11 words per
sentence) than {\it right} endings (8.75 words), which are in turn
slightly longer than {\it wrong} ones (8.47 words). 
Previous work has shown that sentence length is also indicative of whether a text was deceptive \cite{yancheva2013automatic,qin2004exploratory}. 
Although writing {\it wrong} sentences is not the same as lying, it is not entirely surprising to observe similar trends in both tasks.
%Furthermore, \citet{Buller:1996} suggests lying increases one's cognitive load, which in turn could affect sentence length.
%\yc{Is there psycholinguistic literature that we can cite that may have shown that cognitive burden causes writers to write shorter sentences...?}

@@

Finally, \figref{roc_word_distribution} presents the distribution of five frequent words across the different groups. 
The figure shows that {\it original} endings use coordinations (``and'') more than  {\it right} endings, and substantially more than {\it wrong} ones. %\ms{\cite{Newman:2003} found that liars use less ``contrast'' coordination, did we find that in our experiments?}.
Furthermore, {\it original} and {\it right} endings seem to prefer enthusiastic language (e.g., ``!''), while {\it wrong} endings tend to use more negative language (``hates''), similarly as deceptive text \cite{Newman:2003}.
%\yc{Again, any cogsci literature on this?} \nascomment{feels
%  Pennebaker-ish, that might be a good place to look.}
Next we show that these style differences are not anecdotal, but can
be used to distinguish between the different types of story endings.

@@

\paragraph{Story cloze task.}
The results of Experiment 1 indicate that {\it right} and {\it wrong} endings are characterized by different styles.
In order to further estimate the quality of our classification results, we tackle the story cloze task using our classifier.
This classification task is more constrained than Experiment 1, as two
endings are given and the question is which is \emph{right} and which is
\emph{wrong}.
We apply the classifier from Experiment 1 as follows:
if it assigns different labels to the two given endings, we keep
them.  If not, the label whose posterior probability is lower is reversed.

@@

\begin{table}[!t]
\begin{center}
\begin{tabular}{|l|r|} \hline
{\bf Model} & {\bf Acc.} \\ \hline
$\dagger${DSSM} \cite{Mostafazadeh:2016} & 0.585 \\ 
$\dagger${LexVec} \cite{Salle:2016} & 0.599 \\ \hline\hline
%{Niko (shared task)}	& 0.700\\ 
%{tbmihaylov (shared task)} & 0.711\\ \hline\hline
$\dagger${RNN}		& 0.677 \\ 
{\bf Ours} & {\bf 0.724} \\ 
$\dagger${\bf Combined (ours + RNN)} & {\bf 0.752} \\ \hline\hline
$\dagger$Human judgment & 1.000 \\ \hline
\end{tabular}
\end{center}
\caption{\label{cloze_results}\resolved{\yc{we should change accuracy to be in 0-100\% scale to be consistent with Table 2}}Results on the test set of the  story cloze task. 
The first block are published results, 
the second block is our results.
LexVec results are taken from \cite{Speer:2016}.
%RNN is our implementation. 
Human judgement scores are taken from \cite{Mostafazadeh:2016}.
Methods marked with ($\dagger$) use the story context in order to make a prediction. 
 \resolved{\nascomment{I added 0s to ``Niko'' result
  and ``human''
  so sig. digits would line up.  if we don't have sig digits for some
  of them, make the zeros white so the numbers line up.  where did the
human judgments come from?  those are not mentioned in our text anywhere!}
}}
\end{table}

@@

\paragraph{Experiment 2.}
%In order to understand which features were most salient, we repeated Experiment 1, this time running SVM with L1 norm, such that it generates a sparse separating hyperplane. 
%This came at a minor cost of less than 2\% in performance compared to our L2 results (0.697 compared to 0.724).  
\tabref{exp2_features} shows the same analysis for Experiment 2.
As noted in \secref{ROC_Story}, {\it original} endings tend to be much longer, which is indeed the most salient feature for them.
This is in line with similar findings in the deception detection literature \cite{qin2004exploratory}.
An interesting observation is that exclamation marks are a strong
indication for an  {\it original} ending. 
This suggests that authors are more likely to show or evoke enthusiasm when writing their own text compared to ending an existing task.

@@

\paragraph{Writing style.}
Writing style has been an active topic of research for decades. 
The models used to characterize style are often linear classifiers with style features such as character and word $n$-grams \cite{Stamatatos:2009,Koppel:2009}.
Previous work has shown that different authors can be grouped by their
writing style, according to factors such as age
\cite{Pennebaker:2003,Argamon:2003,Schler:2006,Rosenthal:2011,nguyen:2011:latech},
gender \cite{Argamon:2003,Schler:2006,bamman2014gender}, and native language
\cite{Koppel:2005,Tsur:2007,Bergsma:2012}.   \resolved{\nascomment{cite generously here, it costs us nothing}}
At the extreme case, each individual author adopts a unique writing
style \cite{mosteller1963inference,pennebaker1999linguistic,Schwartz:2013}. 
\com{Interestingly, previous work has shown that individual style can be affected from coarse-grained factors such as those just described, but also from other less apparent factors such as mental state, or even living in a high-elevation location \cite{schwartz2013personality}.

@@

The line of work that most resembles our work is the detection of deceptive text. 
Several researchers have used stylometric features to  predict deception 
\cite{Newman:2003,hancock2007lying,ott2011finding,Feng:2012}.
Some works even showed that gender affects a person's  writing style when lying \cite{Perez:2014a,Perez:2014b}.
In this work, we have shown that an even more subtle writing task---writing {coherent} and {incoherent} story endings---imposes different styles on the author.

>>>>

\begin{document}
\maketitle
\begin{abstract}
A writer's style depends not just on personal traits but also on her intent and mental state.
In this paper, we show how variants of the same writing task can lead to measurable differences in writing style.
We present a case study based on 
the  {\it story cloze task} \cite{Mostafazadeh:2016},
where annotators were assigned similar writing tasks with different constraints: (1) writing an entire story, (2) adding a story ending for a given story context, and (3) adding an incoherent ending to a story.\resolved{\clinic{Many felt these 2 constraints looks like 3 constraints. clarify}}
%We show that a linear classifier, which applies only simple style features, such as sentence length and character n-grams, 
We show that a simple linear classifier informed with stylistic features is able to successfully distinguish among the three cases, without even looking at the story context.
In addition, our style-based classifier establishes a new state-of-the-art result on the story cloze challenge, substantially higher than previous results based on deep learning models. 
%Furthermore, combining our model with an LSTM-based language model, which does look at the context, leads to additional gains.
Our results demonstrate that different task framings can dramatically affect the way people write.\resolved{\clinic{Similarly to previous comment: which results does this comment address?}}

@@

\section{Introduction}
Writing style is expressed through a range of linguistic elements such as words, sentence structure, and rhetorical devices.
It is influenced by personal factors such as age and gender \cite{Schler:2006}, 
by personality traits such as agreeableness and openness  \cite{Ireland:2014b},
as well as by mental states\resolved{\clinic{Avoid using ``cognitive"?}} such as sentiment \cite{Davidov:2010}, sarcasm \cite{Tsur:2010}, and deception \cite{Feng:2012}.  
In this paper, we study the extent to which writing style is affected by the nature of the writing task the writer was asked to perform, since
different tasks likely engage different cognitive processes \cite{Campbell:2003,Banerjee:2014}.

@@

While the story cloze task was originally designed to be a story understanding challenge, 
its annotation process introduced three variants of the same writing task: writing an {\it original}, {\it right}, or {\it wrong} ending to a short story.
In this paper, we show that a linear classifier informed by stylistic features can distinguish among the different endings to a large degree, even without looking at the story context (64.5--75.6\% binary classification results).

@@

In order to further validate our method, we also directly tackle the
story cloze task\nascomment{cite the challenge?  a reader may
  not realize that there is a shared task on this}\roy{Maybe ``challenge'' was not the right term here. We basically mean the story cloze task. The shared task was the original task w/o any modification, so I am not referring to it specifically here}. 
 Adapting our classifier to the task, we obtain 72.4\% accuracy, a
 12.5\% increase over the previously reported state-of-the-art  \cite{Salle:2016}.
We also show that the style differences captured by our model can be combined with neural language models to make a better use of the story context. 
Our final model that combines context with stylistic features achieves
75.2\%---an additional 2.8\% gain, 15.3\% better than the best
published result.

@@

The resulting stories, both {\it right} and {\it wrong}, were then individually rated for coherence and meaningfulness by additional AMT workers.
Only stories rated as simultaneously coherent with a {\it right} ending and neutral with a {\it wrong} ending were selected for the task. 
It is worth noting that workers rated the stories as a whole, not only the endings.

@@

\isection{Surface Analysis of the Story Cloze Task}{Surface}
We begin by computing several characteristics of the three types of endings: {\it original} endings (from the ROC story corpus training set), {\it right} endings and {\it wrong} endings (both from the story cloze task development set).
Our analysis  reveals several style differences between different groups. 
First, {\it original} endings are on average longer (11 words per
sentence) than {\it right} endings (8.75 words), which are in turn
slightly longer than {\it wrong} ones (8.47 words). 
Previous work has shown that sentence length is also indicative of whether a text was deceptive \cite{yancheva2013automatic,qin2004exploratory}. 
Although writing {\it wrong} sentences is not the same as lying, it is not entirely surprising to observe similar trends in both tasks.
%Furthermore, \citet{Buller:1996} suggests lying increases one's cognitive load, which in turn could affect sentence length.
%\yc{Is there psycholinguistic literature that we can cite that may have shown that cognitive burden causes writers to write shorter sentences...?}

@@

Finally, \figref{roc_word_distribution} presents the distribution of five frequent words across the different groups. 
The figure shows that {\it original} endings use coordinations (``and'') more than  {\it right} endings, and substantially more than {\it wrong} ones. %\ms{\cite{Newman:2003} found that liars use less ``contrast'' coordination, did we find that in our experiments?}.
Furthermore, {\it original} and {\it right} endings seem to prefer enthusiastic language (e.g., ``!''), while {\it wrong} endings tend to use more negative language (``hates''), similarly as deceptive text \cite{Newman:2003}.
%\yc{Again, any cogsci literature on this?} \nascomment{feels
%  Pennebaker-ish, that might be a good place to look.}
Next we show that these style differences are not anecdotal, but can
be used to distinguish among the different types of story endings.

@@

\paragraph{Story cloze task.}
The results of Experiment 1 indicate that {\it right} and {\it wrong} endings are characterized by different styles.
In order to further estimate the quality of our classification results, we tackle the story cloze task using our classifier.
This classification task is more constrained than Experiment 1, as two
endings are given and the question is which is \emph{right} and which is
\emph{wrong}.
We apply the classifier from Experiment 1 as follows:
if it assigns different labels to the two given endings, we keep
them.  Otherwise, the label whose posterior probability is lower is reversed.

@@

\begin{table}[!t]
\begin{center}
\begin{tabular}{|l|r|} \hline
{\bf Model} & {\bf Acc.} \\ \hline
$\dagger${DSSM} \cite{Mostafazadeh:2016} & 0.585 \\ 
$\dagger${LexVec} \cite{Salle:2016} & 0.599 \\ \hline\hline
%{Niko (shared task)}	& 0.700\\ 
%{tbmihaylov (shared task)} & 0.711\\ \hline\hline
$\dagger${RNN}		& 0.677 \\ 
{\bf Ours} & {\bf 0.724} \\ 
$\dagger${\bf Combined (ours + RNN)} & {\bf 0.752} \\ \hline\hline
$\dagger$Human judgment & 1.000 \\ \hline
\end{tabular}
\end{center}
\caption{\label{cloze_results}\resolved{\yc{we should change accuracy to be in 0-100\% scale to be consistent with Table 2}}Results on the test set of the  story cloze task. 
The upper block shows published results, 
the middle block are our results.
LexVec results are taken from \cite{Speer:2016}.
%RNN is our implementation. 
Human judgement scores are taken from \cite{Mostafazadeh:2016}.
Methods marked with ($\dagger$) use the story context in order to make a prediction. 
 \resolved{\nascomment{I added 0s to ``Niko'' result
  and ``human''
  so sig. digits would line up.  if we don't have sig digits for some
  of them, make the zeros white so the numbers line up.  where did the
human judgments come from?  those are not mentioned in our text anywhere!}
}}
\end{table}

@@

\paragraph{Experiment 2.}
%In order to understand which features were most salient, we repeated Experiment 1, this time running SVM with L1 norm, such that it generates a sparse separating hyperplane. 
%This came at a minor cost of less than 2\% in performance compared to our L2 results (0.697 compared to 0.724).  
\tabref{exp2_features} shows the same analysis for Experiment 2.
As noted in \secref{ROC_Story}, {\it original} endings tend to be much longer, which is indeed the most salient feature for them.
%This is in line with similar findings in the deception detection literature \cite{qin2004exploratory}.
An interesting observation is that exclamation marks are a strong
indication for an  {\it original} ending. 
This suggests that authors are more likely to show or evoke enthusiasm when writing their own text compared to ending an existing task.

@@

\paragraph{Writing style.}
Writing style has been an active topic of research for decades. 
The models used to characterize style are often linear classifiers with style features such as character and word $n$-grams \cite{Stamatatos:2009,Koppel:2009}.
Previous work has shown that different authors can be grouped by their
writing style, according to factors such as age
\cite{Pennebaker:2003,Argamon:2003,Schler:2006,Rosenthal:2011,nguyen:2011:latech},
gender \cite{Argamon:2003,Schler:2006,bamman2014gender}, and native language
\cite{Koppel:2005,Tsur:2007,Bergsma:2012}.
At the extreme case, each individual author adopts a unique writing
style \cite{mosteller1963inference,pennebaker1999linguistic,Schwartz:2013}. 
\com{Interestingly, previous work has shown that individual style can be affected from coarse-grained factors such as those just described, but also from other less apparent factors such as mental state, or even living in a high-elevation location \cite{schwartz2013personality}.

@@

The line of work that most resembles our work is the detection of deceptive text. 
Several researchers have used stylometric features to  predict deception 
\cite{Newman:2003,hancock2007lying,ott2011finding,Feng:2012}.
Some works even showed that gender affects a person's  writing style when lying \cite{Perez:2014b,Perez:2014a}.
In this work, we have shown that an even more subtle writing task---writing {coherent} and {incoherent} story endings---imposes different styles on the author.
